---
title: "Ambrosioni_Hélène_Homework2"
author: "Hélène Ambrosioni"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```
<center>
# **Homework II**
</center>

#### <u>The Enviroment, Library and Data  </u>

We need some libraries
```{r eval=FALSE}
library(lubridate) 
library(readxl) 
library(ggplot2) 
library(plotly)
library(corrplot)
library(dplyr)
library(ggcorrplot)
library(Hmisc)
library(qgraph)
library(readr)
library(car)
library(knitr)
```

```{r include=FALSE}
bitcoin_price <- read.csv("exchanges.csv")
data <- bitcoin_price
```

## <u> 1 Exercise </u>

Answers to questions:

**What period does that data cover?**

```{r echo = FALSE, results='asis'}
cat("- The period covered by the data is between ", min(data$Date), "and", max(data$Date), "\n")
```

**What data types are included in the data set?**

```{r echo = FALSE, results='asis'}
library(skimr)
skim(data)
```

If you want a more "rough" and detailed version of the dataframe structure instead:

```{r echo = FALSE, results='asis'}
str(data)
```

**Now, we want to check and evaluate the quality of the data**
```{r echo = FALSE}
library(knitr)

# Control of missing values for data quality
missing_values <- sum(is.na(data))
missing_values_table <- data.frame(Sum_Missing_Values = missing_values)
kable(missing_values_table, caption = "Total Missing Values")

# Missing values for each column
missing_values_columns <- colSums(is.na(data))
missing_values_columns_table <- data.frame(Column = names(missing_values_columns), Missing_Values = missing_values_columns)
kable(missing_values_columns_table, caption = "Missing Values per Column")
```

* In this dataset, I personally rate the data very good, there is no missing data.  


**What are the mean and median of the numerical variables included?**


```{r echo = FALSE, results='asis'}

numeric_columns <- sapply(data, is.numeric)

for (column in names(data[numeric_columns])) {
  mean_val <- mean(data[[column]], na.rm = TRUE)
  median_val <- median(data[[column]], na.rm = TRUE)
  cat(column, " | Mean:", mean_val, " | Median:", median_val, "\n\n")

}
```


**Which variable has the largest variation in the observed period?**

- Column with maximum deviation:
```{r echo = FALSE, results='asis'}

numeric_columns <- sapply(data, is.numeric)

max_deviation <- -1
col_with_max_deviation <- ""


for (column in names(data[numeric_columns])) {
  deviation <- sd(data[[column]], na.rm = TRUE)
  if (deviation > max_deviation) {
    max_deviation <- deviation
    col_with_max_deviation <- column
  }
}

cat(col_with_max_deviation, "\n\n")
```
- Minimum deviation:
```{r echo=FALSE, results='asis'}
cat( max_deviation, "\n\n")
```


**Which variable has the smallest variation in the observed period?**

- Column with minimum deviation:
```{r echo=FALSE, results='asis'}
# Codice per trovare la deviazione minima
numeric_columns <- sapply(data, is.numeric)

min_deviation <- Inf
col_with_min_deviation <- ""

for (column in names(data[numeric_columns])) {
  deviation <- sd(data[[column]], na.rm = TRUE)
  if (deviation < min_deviation) {
    min_deviation <- deviation
    col_with_min_deviation <- column
  }
}

cat(col_with_min_deviation, "\n\n")
```
- Minimum deviation:
```{r echo=FALSE, results='asis'}
cat( min_deviation, "\n\n")
```

## <u> 2 Exercise </u>

**What is the mean price of the Bitcoin cryptocurrency (exchange: coinbase) for the period before and after 2017?**

- The Mean price of the Bitcoin before 2017 is: 
```{r echo=FALSE, results='asis'}
library(lubridate)
library(dplyr)
# Changing the format of dates
data$Date <- as.Date(data$Date, format = "%m/%d/%Y")

data$Date <- as.Date(data$Date)


# Create the binary variable
data$Period <- ifelse(year(data$Date) < 2017, "Before", "After")
# Selects only the columns 'Date' and 'btc_coinbase'
bitcoin_data <- data[c('Date', 'btc_coinbase')]

# The 'Date' column converses in date format
bitcoin_data$Date <- as.Date(bitcoin_data$Date)

# Add a new 'Period' column based on the date
bitcoin_data <- bitcoin_data %>%
  mutate(Period = ifelse(year(Date) < 2017, "Before", "After"))


# Filtering data for the period before and after 2017 on Coinbase for Bitcoin
before_2017_coinbase <- subset(data, Date < as.Date("2017/01/01") & btc_coinbase > 0)
after_2017_coinbase <- subset(data, Date >= as.Date("2017-01-01") & btc_coinbase > 0)

# Calculate the average price before and after 2017 on Coinbase for Bitcoin
mean_before_2017_coinbase <- mean(before_2017_coinbase$btc_coinbase)
mean_after_2017_coinbase <- mean(after_2017_coinbase$btc_coinbase)

# Results
cat( mean_before_2017_coinbase, "\n")


```

- The Mean price of the Bitcoin after 2017 is: 
```{r echo=FALSE, results='asis'}

cat( mean_after_2017_coinbase, "\n")
```



**On what day was the difference between Bitcoin and Kraken currencies largest?**

-
```{r echo=FALSE, results='asis'}
# Calculate the difference between Bitcoin prices on Coinbase and Kraken
data$Difference <- data$btc_coinbase - data$btc_kraken

# Find the day with the largest difference
day_of_max_difference <- data$Date[which.max(data$Difference)]
# Format the date as dd/mm/yyyy
formatted_date <- format(day_of_max_difference, "%d/%m/%Y")


# Display the day with the largest difference in dd/mm/yyyy format
cat("On day:", formatted_date, "\n")
```


## <u> 3 Exercise </u>

Investigate the price trend of bitcoin through the use of chart plots

- The trend of the Bitcoin Price, x = Date, y = Bitcoin price in dollars
```{r echo=FALSE, results='asis'}
library(ggplot2)
# Date column is in the Date format
data$Date <- as.Date(data$Date)

# Line graph of bitcoin price on Coinbase
ggplot(data, aes(x = Date, y = btc_coinbase)) +
  geom_line(color = "red", size = 1.0) + 
  labs(title = "The trend of the Bitcoin Price",
       x = "Date",
       y = "Bitcoin price in dollars") +
  theme_minimal()

```

- Histogram of the Bitcoin price

```{r echo=FALSE, results='asis'}
ggplot(data, aes(x = btc_coinbase)) +
  geom_histogram(binwidth = 500, fill = "magenta", color = "black") + 
  labs(title = "Bitcoin price distribution on Coinbase",
       x = "Bitcoin price in dollars",
       y = "Frequency") +
  theme_minimal()
```

- Box plot of the Bitcoin price

```{r echo=FALSE, results='asis'}
ggplot(data, aes(y = btc_coinbase)) +
  geom_boxplot(fill = "lightpink", color = "black") + 
  labs(title = " Bitcoin price distribution on Coinbase ",
       y = "Bitcoin price in dollars") +
  theme_minimal()
```

- Boxplot on Bitcoin more accurate

```{r echo=FALSE, results='asis'}
ggplot(data, aes(y = btc_coinbase)) +
  geom_boxplot(fill = "lightpink", color = "black") + 
  labs(title = " Bitcoin price distribution on Coinbase ",
       y = "Bitcoin price in dollars") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 5000))
```


**What can be said about the price of Bitcoin over the observed period?**

## <u> 4 Exercise </u>

Investigate to the relationship between the price of Bitcoin and that of traditional assets

- The dependence between the Bitcoin price and the price of oil
```{r echo=FALSE, results='asis'}
# Plot and regression line plotting - Bitcoin and oil relationship
ggplot(data, aes(x = oil, y = btc_coinbase)) +
  geom_point(color = "blue") +  
  labs(title = "Relationship between the price of Bitcoin and the price of oil",
       x = "Oil price",
       y = "Bitcoin price on Coinbase") +
  theme_minimal() +
  geom_smooth(method = "lm", se = FALSE, color = "red")  
```

**What can you tell about the dependency between these two assets?**

- The dependence between the Bitcoin price and the price of gold
```{r echo=FALSE, results='asis'}
#Correct columns
bitcoin_gold_data <- data[c('btc_coinbase', 'gold')]

# #Plot and regression line plotting - Bitcoin and gold relationship
ggplot(bitcoin_gold_data, aes(x = gold, y = btc_coinbase)) +
  geom_point(color = "black") +  
  labs(title = "Dependence between the price of Bitcoin and the price of gold",
       x = "Gold price",
       y = "Bitcoin price on Coinbase") +
  theme_minimal() +
  geom_smooth(method = "lm", se = FALSE, color = "magenta")  

```


**What can you tell about the dependency between these two assets?**



Predict the price of Bitcoin (exchange = coinbase) using the price of oil, gold, and the SP500.
```{r echo=FALSE, results='asis'}
library(plotly)
#Correct columns
bitcoin_gold_data <- data[c('btc_coinbase', 'gold')]

# Plotly
plot_ly(data = bitcoin_gold_data, x = ~gold, y = ~btc_coinbase, type = 'scatter', mode = 'markers', marker = list(color = 'blue', size = 3)) %>%
  layout(title = "Dependence between the price of Bitcoin and the price of gold",
         xaxis = list(title = "Gold price"),
         yaxis = list(title = "Bitcoin price on Coinbase"),
         showlegend = FALSE) %>%
  add_lines(x = ~gold, y = ~fitted(lm(btc_coinbase ~ gold, data = bitcoin_gold_data)), line = list(color = 'red'), inherit = FALSE)


# Multiple regression model using the price of oil, gold and the SP500
model <- lm(btc_coinbase ~ oil + gold + sp500, data = data)

```

-  Multiple regression model

```{r echo=FALSE, results='asis'}
library(car)
# Multiple regression model
model <- lm(btc_coinbase ~ oil + gold + sp500, data = data)

# Plotting avPlots graphs for independent variables
avPlots(model)
```

## <u> 5 Exercise </u>

- Correlations between all numeric data in the data set

```{r echo=FALSE, results='asis'}
# Select only numeric columns.
numeric_columns <- data[, sapply(data, is.numeric)]

# Calculate the correlation matrix
corr <- cor(numeric_columns, use = "complete.obs")
numeric_columns <- data[, sapply(data, is.numeric)]

```



- Visualization of the lower half of the correlation matrix (method = "circle"). 

```{r echo=FALSE, results='asis'}
library(ggcorrplot)
#Lower half of the correlation matrix
ggcorrplot(corr, type = "lower", method = "circle")

```


- Heatmap of the correlations that emerge between the numerical variables included in the data set.

```{r echo=FALSE, results='asis'}
# Calculate correlations between numerical variables
numeric_columns <- data[, sapply(data, is.numeric)]
corr <- cor(numeric_columns, use = "complete.obs")

# Interactive heat map with plotly
plot_ly(z = corr, type = "heatmap") %>%
  layout(title = "Heat map of correlations between numerical variables",
         x = colnames(corr), y = colnames(corr),
         xgap = 1, ygap = 1)

```